\paperTitle

### 2014-05-17 doc/icml2014/directed.tex
simplify directed grid example

\section{Recovering hidden marginals}
\label{sec:exclusiveViews}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example: directed grid model}
\label{sec:directedExample}

To gain some intuition, consider the directed grid model from \figureref{approach}.
%The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  %hidden variables $h_1, \ldots, h_4$.
The parameters of this model are the conditional probability tables
$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$.

\paragraph{Estimating $O$}
Since the observed variables $x^a_1, x^b_1, x^a_2$ are
  conditionally independent given $h_1$, we can $\TensorFactorize$ from
  \sectionref{setup} to recover $O$.

\paragraph{Estimating $\pi$}
The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
  $\pi$ by a linear system: $\mO_1 = O \pi$. 
If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.

\paragraph{Estimating $T$}
Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
  \eqdef \Pr(x^a_1, x^a_2)$, in terms of the hidden marginals $\mH_{12}
  \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$:
\begin{align*}
\mO_{12} = O \mH_{12} O^\top \quad\Rightarrow\quad
  \mH_{12} = O\pinv  \mO_{12} O\pinvt .
\end{align*}
$T$ can be recovered from the $\mH_{12}$ by renormalizing the columns.

\paragraph{Estimating $V$}
Finally, we can estimate $V$ by renormalizing the hidden marginals
$\mH_{234} \eqdef \Pr(h_2, h_3, h_4)$ from the third-order moments
$\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$:
\begin{align*}
  \mO_{234} = \mH_{234}(O, O, O) \quad\Rightarrow\quad
  \mH_{234} = \mO_{234}(O\pinv , O\pinv , O\pinv ).
\end{align*}

@@

The intuitions of the directed grid generalize readily to general graphs.
The key property required for our algorithm to work is as follows:
\begin{definition}[Exclusive views]
  \label{def:exclusive-views}
Let $S \subseteq \bh$ be a subset of hidden variables.
We say $h_i \in S$ has an exclusive view $x_v$
  if the two conditions hold:
  (i) there exists some observed variable
  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$,
  and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ have full column rank $k$ and can be recovered.
We say that $S$ has the \emph{exclusive views property} if every $h_i \in S$ has an exclusive view.
%If every hidden clique has the exclusive views property, then we say that $\sG$ does as well.
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Estimating hidden subset marginals}

>>>>

\section{Recovering hidden marginals}
\label{sec:hiddenMarginals}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Example: directed grid model}
%\label{sec:directedExample}

\paragraph{Example}

%The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  %hidden variables $h_1, \ldots, h_4$.
%The parameters of this model are the conditional probability tables
%$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
%V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
%=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$.

%\paragraph{Estimating $O$}
%Since the observed variables $x^a_1, x^b_1, x^a_2$ are
%  conditionally independent given $h_1$, we can $\TensorFactorize$ from
%  \sectionref{setup} to recover $O$.

%\paragraph{Estimating $\pi$}
%The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
%  $\pi$ by a linear system: $\mO_1 = O \pi$. 
%If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.

%\paragraph{Estimating $T$}

To gain some intuition, consider the directed grid model from \figureref{approach}.
We can express the observed marginals $\mO_{12} \eqdef \Pr(x^a_1, x^a_2) \in \Re^{d \times d}$
as a linear function of the hidden marginals $\mH_{12} \eqdef \Pr(h_1, h_2) \in \Re^{k \times k}$,
where the linear coefficients are based on the conditional moments
$\mOpp{1}{1},\mOpp{2}{2} \in \Re^{d \times k}$:
%down the moments of $x^a_1, x^a_2$,
\begin{align*}
\mO_{12} = \mOpp{1}{1} \mH_{12} \mOppt{2}{2}.
\end{align*}
We can then solve for $\mH_{12}$ by matrix inversion:
\begin{align*}
\mH_{12} = \mOppi{1}{1} \mO_{12} \mOppit{2}{2}.
\end{align*}
%$T$ can be recovered from the $\mH_{12}$ by renormalizing the columns.

%\paragraph{Estimating $V$}
%Finally, we can estimate $V$ by renormalizing the hidden marginals
%$\mH_{234} \eqdef \Pr(h_2, h_3, h_4)$ from the third-order moments
%$\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$:
%\begin{align*}
%  \mO_{234} = \mH_{234}(O, O, O) \quad\Rightarrow\quad
%  \mH_{234} = \mO_{234}(O\pinv , O\pinv , O\pinv ).
%\end{align*}

@@

%The intuitions of the directed grid generalize readily to general graphs.
%The key property required for our algorithm to work is as follows:
For which subsets of hidden nodes can we recover the marginals?  The following
definition offers an characterization:
\begin{definition}[Exclusive views]
  \label{def:exclusive-views}
Let $S \subseteq \bh$ be a subset of hidden variables.
We say $h_i \in S$ has an exclusive view $x_v$
  if the two conditions hold:
  (i) there exists some observed variable
  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$,
  and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ have full column rank $k$ and can be recovered.
We say that $S$ has the \emph{exclusive views property} if every $h_i \in S$ has an exclusive view.
%If every hidden clique has the exclusive views property, then we say that $\sG$ does as well.
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Estimating hidden marginals}

### 2014-05-17 doc/icml2014/intro.tex
simplify directed grid example

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals,
% then get parameters.
% DONE: try to talk about the approach in terms of three steps
The essence of our approach is illustrated in \figureref{approach},
which contains three steps.
First, we identify three views for each hidden variable $h_i$ (for example,
$x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
the tensor factorization algorithm of
\citet{anandkumar13tensor} to estimate the \emph{conditional
moments} $\Pr(x_i^a \mid h_i)$ and $\Pr(x_i^b \mid h_i)$ for each $i$ (\sectionref{bottlenecks}).
Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
subsets of hidden nodes (e.g., $\mathbb P(h_2, h_3, h_4)$).
Normally, such a marginal likelihood objective would be non-convex,
but given the conditional moments, we obtain a convex objective,
which can be globally optimized using EM
(see Sections~\ref{sec:exclusiveViews} and \ref{sec:piecewise}).
So far, our method has relied only on the conditional independence
structure of the model and applies generically to both directed
and undirected models.
The final step of turning hidden marginals into model parameters
requires some specialization.
In the directed case, this is simple normalization;
in the undirected case, we need to solve another convex optimization problem
(\sectionref{undirected}).

>>>>

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals,
% then get parameters.
% DONE: try to talk about the approach in terms of three steps
The essence of our approach is illustrated in \figureref{approach},
which contains three steps.
First, we identify three views for each hidden variable $h_i$ (for example,
$x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
the tensor factorization algorithm of
\citet{anandkumar13tensor} to estimate the \emph{conditional
moments} $\Pr(x_i^a \mid h_i)$ and $\Pr(x_i^b \mid h_i)$ for each $i$ (\sectionref{bottlenecks}).
Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
subsets of hidden nodes (e.g., $\mathbb P(h_2, h_3, h_4)$).
Normally, such a marginal likelihood objective would be non-convex,
but given the conditional moments, we obtain a convex objective,
which can be globally optimized using EM
(see Sections~\ref{sec:hiddenMarginals} and \ref{sec:piecewise}).
So far, our method has relied only on the conditional independence
structure of the model and applies generically to both directed
and undirected models.
The final step of turning hidden marginals into model parameters
requires some specialization.
In the directed case, this is simple normalization;
in the undirected case, we need to solve another convex optimization problem
(\sectionref{undirected}).
